PeerMind

### 2016-08-11 paper.tex
Paragraphs.
Normalize delegations.

%\pagenumbering{arabic}

@@

How will you know how well you solved the problem?
If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
So currently we have issues with reaching the quorum and many people do not participate.
I would like for them to have easier time participating (less time required) and as a consequence more people to
participate collectively.
If I observe that I think I will know I solved the problem.

@@

\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}

@@

\begin{itemize}
\item Member $M_0$ decides not to delegate at all.
\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
\item Member $M_3$ delegates to $M_1$ and $M_2$ with ratios $0.40$ ($r_{3,1}$) and $0.40$ ($r_{3,2}$).
\end{itemize}

Observe that member $M_3$ decided to use only $0.8$ of their vote in the case they do not vote themselves.
The rest ($0.2$) counts towards non-voting population.
In this way a member can decide to participate in voting through delegation, but not fully.
Member $M_0$ decided to not participate at all and if they do not vote they are counted towards non-voting population.

@@

    \Vertex[Math,L=M_0]{M0}
    \EA[Math,L=M_1](M0){M1}
    \SO[Math,L=M_3](M1){M3}
    \WE[Math,L=M_2](M3){M2}

    \Edge[label=$0.40$](M1)(M3)
    \Edge[label=$0.75$](M0)(M1)
    \Edge[label=$0.25$](M3)(M1)

@@

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & 0 \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} = 0.40 V_{M_1} + 0.40 V_{M_2} \\
\end{array}
\end{displaymath}

@@

\begin{displaymath}
\mathrm{D} = \left[ \begin{array}{cccc}
1 & 0 & 0 & 0 \\
-0.75 & 1 & -0.25 & 0 \\
0 & 0 & 1 & 0 \\
0 & -0.40 & -0.40 & 1 \\
\end{array} \right]
\end{displaymath}

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.7 \\
0 \\
-0.4 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.7 \\
0.54 \\
-0.4 \\
0.056 \\
\end{array} \right]
\end{displaymath}

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.22$.
Delegation pushes decision slightly in favor.

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded and should count
towards non-voting population.
Members might not declare their delegates, or they might declare them without using their vote fully.
Other members can delegate to them as well, potentially propagating lack of a vote further.
For example, if member $M_0$ would not vote, then their vote would be discarded.
Moreover, we could not use $M_0$'s vote to compute $M_1$'s vote fully, which would influence $M_3$'s vote as well.

One option is that we would normalize ratio in such case, so if $M_0$ does not vote, then $r_{1,3}$ is normalized
from $0.25$ to $1.0$.
We could do this automatically for all members, or we could even leave to members to configure
what to do in such case for them.
In current version of PeerMind we decided to not handle this situation in any special way and we leave
discarded votes to propagate.

\todo[inline]{Or should we be normalizing? In a way we should not normalize if votes are discarded because
member explicitly delegated so ($M_3$ above), but if votes are discarded because we could not compute
all delegations, we might want to automatically normalize.}

As a consequence we have to determine how many votes have been discarded to know the size of the non-voting
population.
Another issue is also a question of abstentions.
In PeerMind members can abstain from voting which is different from simply not participating.
Abstentions are counted towards quorum, but they do not contribute to the result.
To address this, we define two vectors:

@@

Solving $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ gives
us number of votes $W$ and number of abstentions $A$, respectively, by summing together all values in the
resulting vectors: $W = \sum \mathbf{x_w}$, $A = \sum \mathbf{x_a}$.

\todo[inline]{Is the result then $\frac{\sum \mathbf{x_v}}{W}$ or $\frac{\sum \mathbf{x_v}}{\textrm{number of entries
in}~\mathbf{x_w}~\textrm{which are non-zero}}$?}

\todo[inline]{We also have to explain that with non-zero values of $\mathbf{x_w}$ we also get which zero values in
$\mathbf{x_v}$ are real results and which are zero because delegation could not be computed. One has to use that
to compute delegated votes as well.}

We pass computed votes and number of votes and abstentions into computation of dynamic quorum to determine
a quorum and confidence, in the same manner as if members voted themselves.

>>>>

\pagenumbering{arabic}

@@

How will you know how well you solved the problem?

If it improves decision making, participation, inclusivity, activity of people engaged in decision making in my house.
So currently we have issues with reaching the quorum and many people do not participate.
I would like for them to have easier time participating (less time required) and as a consequence more people to
participate collectively.
If I observe that I think I will know I solved the problem.

@@

\todo[inline]{Terminology: decision vs. vote vs. voting (how to name the step in decision making where you vote, and what is what you cast?)}
\todo[inline]{Terminology: motion? Tally? What is the sum of all votes?}
\todo[inline]{Terminology: discarding a vote if you didn't vote and didn't delegate. Some other term?}

@@

\begin{itemize}
\item Member $M_0$ decides not to delegate at all.
\item Member $M_1$ delegates to $M_0$ and $M_3$ with ratios $0.75$ ($r_{1,0}$) and $0.25$ ($r_{1,3}$), respectively.
\item Member $M_2$ delegates to $M_3$ with ratio $1.00$ ($r_{2,3}$).
\item Member $M_3$ delegates to $M_1$, $M_2$, and $M_4$ with ratios $0.40$ ($r_{3,1}$), $0.40$ ($r_{3,2}$), and $0.20$ ($r_{3,4}$), respectively.
\item Member $M_4$ decides not to delegate.
\end{itemize}

The sum of ratios of all delegations of a member has to be $1.0$.
Members $M_0$ and $M_4$ decided to not participate in delegation and if they do not vote they are
counted towards non-voting population.

@@

    \Vertex[Math,L=M_0]{M0}
    \EA[Math,L=M_1](M0){M1}
    \SO[Math,L=M_3](M1){M3}
    \WE[Math,L=M_2](M3){M2}
    \EA[Math,L=M_4](M3){M4}

    \Edge[label=$0.40$](M1)(M3)
    \Edge[label=$0.75$](M0)(M1)
    \Edge[label=$0.25$](M3)(M1)
    \Edge[label=$0.20$](M4)(M3)

@@

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & 0 \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
\end{array}
\end{displaymath}

@@

\begin{displaymath}
\mathrm{D} = \left[ \begin{array}{ccccc}
1 & 0 & 0 & 0 & 0 \\
-0.75 & 1 & -0.25 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 \\
0 & -0.40 & -0.40 & 1 & -0.20 \\
0 & 0 & 0 & 0 & 1 \\
\end{array} \right]
\end{displaymath}

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.7 \\
0 \\
-0.4 \\
0 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.7 \\
0.54 \\
-0.4 \\
0.056 \\
0.0 \\
\end{array} \right]
\end{displaymath}

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded and should count
towards non-voting population.
Members might not declare their delegates.
Other members can delegate to them as well, potentially propagating lack of a vote further.
In our example, member $M_4$ has not voted so their vote is discarded.
As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.

In such case we normalize ratios of other delegates, those who voted or for which we can
compute their votes recursively.
So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.

Another issue is a question of abstentions.
In PeerMind members can abstain from voting which is different from simply not participating.
Abstentions are counted towards quorum, but they do not contribute to the result.
To address this, we define two vectors:

@@

We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$.
To normalize votes we compute $x = x_v / x_w$.
A consequence is that those elements of $x_w$ equal to $0$ make corresponding elements of $x$ undefined.
Those undefined elements of $x$ we count towards non-voting population.
We just have to make sure that they are not in fact abstentions.
We check corresponding elements of $x_a$ and if they are non-zero we count them as abstentions,
otherwise as non-votes.
In our example:

\begin{displaymath}
\mathbf{x} = \left[ \begin{array}{c}
0.7 \\
0.57 \\
-0.4 \\
0.07 \\
\mathrm{undefined} \\
\end{array} \right]
\end{displaymath}

\todo[inline]{Are we discarding abstentions too much here? If we have 5 people and each of them is computed 0.2 of abstention
through delegation, while the rest 0.8 is a vote, then we normalize 0.8 to 1.0 for all of them and simply discard 0.2
of abstention. But otherwise those 0.2 of abstentions would count together towards 1.0. But probably this is not a
problem because the quorum is still the same: those 5 people have participated through delegation in voting. Only
if somebody has not participated through delegation in voting we care if they might participated in abstentions
through voting. To maximize delegated participation.}

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.24$.
Delegation pushes decision slightly in favor.

\todo[inline]{$x = x_v / x_w$ does not really produce the same result as computing with 0.50 as ratio to begin
with. Why?}

We pass computed votes and number of abstentions into computation of dynamic quorum to determine
a quorum and confidence, in the same manner as if members voted themselves.
